Grade Retention and Mixed Race

### 2022-08-24 products/paper/main.qmd
Add outline for a new approach to first part of background section

>>>>

Possible restructuring of this whole section:

* To understand where biracials fit, it is helpful to think about what mechanisms drive the racial hierarchy more broadly.
  * First, racial inequality in outcomes may be produced by the cumulative sum of individual acts of discrimination as well as systematic forms of bias and discrimination. Individual discrimination need not be consciously performed and may instead reflect cognitive biases and schema that individuals unknowingly rely upon that produce differential outcomes by race. This discrimination will be more consequential when performed by important gatekeepers of opportunity such as teachers, health care workers, and employers. 
  * Systematic discrimination, on the other hand, reflect biases built into institutional systems that channel rewards to certain racial groups. This can operate through such things as norms, laws, and organizational practices. In our current "colorblind" era, such mechanisms may be formally race neutral, but produce disparate impacts by race. 
  * In either case, the experiences of biracial individuals within such a system of discrimination depend on how they are classified by others. To the extent that biracial individuals are classified as more like one of their monoracial constituent groups. they will receive the same treatment and either reap similar rewards or suffer similar discrimination and penalties.
  * discussion of hypodescent/hyperdescent as governing classification into single-race groups. 
  * Iversen et. al.'s critique of hypodescent/hyperdescent, higlighting:
    * historical specificy of hypodescent and hyperdescent pattern
    * equally consistent with a "succession/recession" perspective
    * regimes governing certain groups may differ due to historical circumstances.
  * How does this questions of classification address the case where biracial individuals might be in-between? A couple of possibilities.
    * Racial ambiguity could lead to more inconsistent classification into single race categories across observers submitting biracials to differing exposure to advantages/distadvantages somewhere between  their two monoracial constituent groups.
    * Greater physical ambiguity may allow biracials to avoid worst aspects of discrimination because they don't match the "ideal type" for most discriminated group. We already know that this process exists among monoracial blacks in terms of skin tone, so this is not unique to multiracials.
    * It is also possible that biracials are "correctly" classified by observers into a "middleman" minority position. There is evidence of this kind of practice in Latin America, and it may become more common in the US (Bonilla-Silva). It is unlikely that observers are capable of differentiating fine divisions of ancestry but rather can put people into a nebulous middleman position.
* Some racial inequality may also be explained by differential distributions of resources across racial groups. On average, non-whites come from a lower SES background than whites and this may indirectly lead to observed racial differences. Although much of the focus of the "race vs. class" debate is on SES characteristics, less concretely defined social and cultural capital differences across race may exacerbate racial inequality.
  * The role of these resource differences in biracial's position depends on what resources the typical parent in an interracial union brings with them. This is a complex and not-well understood process even for easily measurable characteristics like education. ...
  * the black box of symbolic resources.
* Aside from these major mechanisms for producing racial inequality, the experience of hybridity itself may affect the outcomes of biracial individuals relative to their monoracial peers.
  * Marginal man theory
  * Enlarged toolkits
